### 2017-09-15 files/paper/tex/abstract.tex

\begin{abstract}
Autonomous underwater vehicles (AUVs) rely on a variety of sensors -- acoustic, inertial and visual -- for intelligent decision 
making. Due
to its non-intrusive, passive nature, and high information content, vision is an attractive sensing modality, particularly at 
shallower
depths. However, factors such as light refraction and absorption, suspended particles in the water, and color distortion affect 
the quality
of visual data, resulting in noisy and distorted images. AUVs that rely on visual sensing thus face difficult challenges, and 
consequently
exhibit poor performance on vision-driven tasks. This paper proposes a method to improve the quality of visual underwater scenes 
using
Generative Adversarial Networks (GANs), with the goal of improving input to vision-driven behaviors further down the autonomy
pipeline. Furthermore, we show how recently proposed methods are able to generate a dataset for the purpose of such underwater 
image
reconstruction\starnote{Reconstruction meaning what we are doing here, right? So this is a contribution too, to show how CycleGAN 
is used to
generate the dataset, correct?}. For any visually-guided underwater robots, this improvement can result in increased safety and 
reliability
through robust visual perception. To that effect, we present quantitative and qualitative data which demonstrates that images 
corrected
through the proposed approach generate more visually appealing images, and also provide increased accuracy for an underwater diver 
tracking
algorithm.
\end{abstract}