### 2017-09-16 files/paper/tex/conclusions.tex
Minor edits.

\section{Conclusion}

Future work: Getting a better and larger dataset. I think that would help a great deal. Also data augmentation
after CycleGAN, such as gaussian blurring, noise, particle effect, more lighting effects etc.

\section*{Acknowledgment}
The authors are grateful to Oliver Hennigh for his implementation of the Gradient Difference Loss measure.