### 2017-09-25 files/paper/PlainPaper.tex
plain version added

\documentclass[11pt, twocolumn]{article}

\usepackage{graphicx}
\graphicspath{ {images/}{images/figure1/}{images/figure2/}{images/figure3/} {images/edges/}}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{adjustbox}
\usepackage{url}
\usepackage{hyperref} % JS: Remove this if you cannot compile.
\usepackage{multirow}

\pdfminorversion=4
% Junaed's definition of a sidenote to look into
\newcommand{\marginlabel}[1]{\mbox{}\marginpar[\raggedleft\hspace{0pt}{#1}]{
\raggedright\hspace{0pt}{#1}}}
\newcommand{\starnote}[1]{\marginlabel{$\bigstar$}\textbf{[#1]}}
\newcommand{\nostarnote}[1]{}
\newcommand{\baad}[1]{} 
% End Junaed's definition

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\title{Enhancing Underwater Imagery using Generative Adversarial Networks}
\author{Cameron Fabbri$^{1}$, Md Jahidul Islam$^{2}$, and Junaed Sattar$^{3}$
\thanks{				
  The authors are with the Department of Computer Science and Engineering, University of Minnesota-Twin  Cities, MN 55455, USA. \newline
 {\tt\small\{$^{1}$fabbr013, $^{2}$islam034, $^{3}$junaed\}@umn.edu}
}
}

\begin{document}
\maketitle

\input{tex/abstract}
\input{tex_plain_version/introduction}
\input{tex/related}
\input{tex_plain_version/methodology.tex}
\input{tex_plain_version/experiments}
\input{tex/conclusions}

\bibliography{cambibs}
\bibliographystyle{plain}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: